Article figures

### 2021-08-24 src/article_figures.Rmd
Tried new economic parameters, and other changes.
Order variables by regression coefficient.

Extracted some functions to an R script.

  #print(combined_variables)

@@

#all_ids <- c(ids, dummy_ids)

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
  # Note that recode stupidly uses order "old"="new"
  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
  return(s)
}

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
  if (FALSE) {
    # df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
    # df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
    # df <- bind_rows(df1, df2)
    # #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
    # 
    # df <- df %>%
    #   mutate(Id=factor(Id, levels=rev(all_ids)),
    #          type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
    #          sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
    #          model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
    #   mutate(type = fct_recode(type, !!!new_names))
    # 
    # thresholds <- get_threshold_df()
    # df <- bind_rows(df, thresholds)
    # df <- df %>% 
    #   mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
    #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
  } else {
#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
    df <- read_tsv(all_results_filename)

    df <- df %>% 
      filter(variable %in% c("AUROC", "AUPR", "F1", "E6", "E12", "threshold6", "threshold12")) %>%
      pivot_wider(names_from=c("type"), values_from=value) %>%
      mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
             high = ifelse(is.na(high), value, high))

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=case_when(variable == "threshold6" ~ "t6",
                             variable == "threshold12" ~ "t12",
                             TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "baseline", "t6", "t12")),
             variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
             variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
             variable = fct_recode(variable, !!!new_names))

  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(variable %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, variable=mytype) # Dummy table to use the same x-axis limits for E6 and E12

  dummy <- dummy %>%
    mutate(variable = fct_recode(variable, !!!new_names))
  #print(dummy)
  #print(df)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and model", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "baseline"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Baseline"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("variable", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  return(list(g=g, df=df))
}

```{r}
res <- create_performance_forest_plot()
g <- res$g
df <- res$df
if (save_figs) {
  filename <- sprintf("%s/performance-forest-plot.pdf", fig_path)
  ggsave(filename=filename, #  title="Performance forest plot", 
         plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
}
g
```

@@

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
    df <- df %>% 
      mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
    prediction_threshold <- hb_threshold
  } else { # predict deferral based on probability of deferral
    df <- df %>% 
      mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
    prediction_threshold <- min(df %>% filter(scores <= probability_of_deferral_threshold) %>% pull(predicted))
  }

@@

  # Get the R2 correlation
  r2 <- summary(lm(predicted ~ observed, data=df))$r.squared
  r2_text <- sprintf("%.2f", r2)

  df <- df %>%
    mutate(confusion_class = factor(ifelse(deferral == 1, 
                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))

@@

  cat(sprintf("Incorrectly predicting %.1f%% of accepted donations as deferrals\n", 100*fp/accepted))
  cat(sprintf("Correctly predicting %.1f%% of deferred donations as deferrals\n", 100*tp/deferrals))

  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_colour_discrete(drop=FALSE) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = prediction_threshold, linetype = "dashed") +
    annotate("text", x=120, y=190, color="black", label=bquote(R^2==.(r2_text)), fontface="bold") +
    theme(legend.position = "bottom")
    #ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")

```{r}
g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.04)
if (save_figs) {
  ggsave(filename="classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g1, dpi=600, units="mm", width=180)
  ggsave(filename="classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g2, dpi=600, units="mm", width=180)
}
g1
g2
```

>>>>

  # Get pretty variable names in the order of female regression coefficients and secondarily by the importances
  get_ordered_levels <- function(df) {
    df <- df %>%
      filter(gender %in% c("female", "both")) %>%
      pivot_wider(id_cols=Pretty, names_from = method, values_from = mean) %>%
      arrange(desc(`Regression coefficient`), desc("Importance"))
    print(df)
    v <- unique(as.character(df$Pretty))
    return(v)
  }

  # Get pretty variable names in the order of average regression coefficient over sex
  get_ordered_levels2 <- function(df) {
    df <- df %>%
      #filter(method == "Regression coefficient") %>%
      group_by(Pretty, method) %>%
      summarise(mean=mean(mean), .groups="keep") %>%
      ungroup() %>%
      pivot_wider(id_cols=Pretty, names_from = method, values_from = mean) %>%
      #mutate(`Regression coefficient`=replace_na(`Regression coefficient`, -1000)) %>%
      arrange(desc(`Regression coefficient`), desc(Importance))
    print(df)
    v <- unique(as.character(df$Pretty))
    return(v)
  }
  #print(combined_variables)

  # Order the variables
  new_levels <- get_ordered_levels2(final)
  new_levels <- rev(new_levels)
  print(new_levels)
  final <- final %>% mutate(Pretty = fct_relevel(Pretty, new_levels))

@@

source("article_figures_functions.R")
```

```{r}
df <- read_tsv(all_results_filename)
res <- create_performance_forest_plot(df)
g <- res$g
df <- res$df
if (save_figs) {
  filename <- sprintf("%s/performance-forest-plot.pdf", fig_path)
  ggsave(filename=filename, #  title="Performance forest plot", 
         plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
}
g
```

@@

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
    hlinetype <- "dashed"
    df <- df %>% 
      mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
    prediction_threshold <- hb_threshold
  } else { # predict deferral based on probability of deferral
    cat(sprintf("Probability threshold is %.2f\n", probability_of_deferral_threshold))
    hlinetype <- "dotted"
    df <- df %>% 
      mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
    prediction_threshold <- min(df %>% filter(scores <= probability_of_deferral_threshold) %>% pull(predicted))
  }

@@

  # Get the R2 correlation
  r2 <- summary(lm(predicted ~ observed, data=df))$r.squared
  #print(r2)
  r2_text <- sprintf("%.2f", r2)
  cat(sprintf("R2 value is %s\n", r2_text))

  df <- df %>%
    mutate(confusion_class = factor(
      case_when(
        deferral == 1 & new_predicted_label == 1 ~ "True positive",
        deferral == 1 & new_predicted_label == 0 ~ "False negative",
        deferral == 0 & new_predicted_label == 1 ~ "False positive",
        deferral == 0 & new_predicted_label == 0 ~ "True negative"
      ),
      # ifelse(deferral == 1, 
      #        ifelse(new_predicted_label == 1, "True positive", "False negative"),
      #        ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))

@@

  cat(sprintf("Incorrectly predicting %.1f%% of accepted donations as deferrals (FPR)\n", 100*fp/accepted))
  cat(sprintf("Correctly predicting %.1f%% of deferred donations as deferrals (TPR)\n", 100*tp/deferrals))

  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_colour_discrete(drop=FALSE) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = prediction_threshold, linetype = hlinetype) +
    annotate("text", x=xymin+10, y=xymax-10, color="black", label=bquote(R^2==.(r2_text)), fontface="bold") +
    theme(legend.position = "bottom")
    #ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
```

```{r Male scatter plot}
g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.08)
if (save_figs) {
  ggsave(filename="classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g1, dpi=600, units="mm", width=180)
  ggsave(filename="classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g2, dpi=600, units="mm", width=180)
}
g1
g2
```

```{r Female scatter plot}
f1 <- create_classification_scatter_plot(finngenn_female_raw$comp_df, 125, NULL)
f2 <- create_classification_scatter_plot(finngenn_female_raw$comp_df, 125, 0.02)
if (save_figs) {
  ggsave(filename="classification-scatter-plot-hb-finngen-female-dlmm.pdf", title="Classifiction scatter plot finngen-female-dlmm", path=fig_path, 
         plot=f1, dpi=600, units="mm", width=180)
  ggsave(filename="classification-scatter-plot-probability-finngen-female-dlmm.pdf", title="Classifiction scatter plot finngen-female-dlmm", path=fig_path, 
         plot=f2, dpi=600, units="mm", width=180)
}
f1
f2
```
